%%=============================================================================
%% Voorwoord
%%=============================================================================

\chapter*{\IfLanguageName{dutch}{Woord vooraf}{Preface}}%
\label{ch:voorwoord}

%% TODO:
%% Het voorwoord is het enige deel van de bachelorproef waar je vanuit je
%% eigen standpunt (``ik-vorm'') mag schrijven. Je kan hier bv. motiveren
%% waarom jij het onderwerp wil bespreken.
%% Vergeet ook niet te bedanken wie je geholpen/gesteund/... heeft

I can proudily say that I have successfully completed my bachelor's thesis. This thesis is the result of hard work, dedication, and perseverance. I am happy to have had the opportunity to conduct this research on two subjects that I am passionate about: Continuous Integration and Deployment (CI/CD) and sustainability.
I would like to express my gratitude to my supervisor, Chantal Teerlinck, for her guidance, support, and valuable feedback throughout the research process. She pushed me through a period of busyness and I will not forget her saying "You have to pass your thesis" all the time. I would also like to thank my internship company, Wolters Kluwer, and mainly my co-supervisor, Jan Delamper, for providing me with the opportunity to conduct this research and for their support and encouragement. 
I would like to extend my heartfelt thanks to my friends and peers for their unwavering support and encouragement. Last but not least, I would like to thank my father with improving my languague and to proofread my thesis. 
I hope that this research will contribute to the field of CI/CD and sustainability and inspire others to explore this topic further.